An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-12-02 paper/3.minbigul_m.tex
add labels to sections
small fixes
Polished text

\section{Adding Memoization: minBiGUL$_m$}

@@

When evaluating a composition of several programs in minBiGUL (and the same holds for full BiGUL), the same $get$s have to be evaluated repeatedly because no intermediate states are kept during the evaluation. This problem was illustrated in Figure~\ref{fig:eval-comp-phead}. To avoid reevaluating $get$s, and as our first approach, we introduce memoization in the minBiGUL interpreter.
% can save 
To keep it simple, the intermediate states of a composition are saved in a key-value table where the key is a pair of program $bx$ and source $s$, and the value is the result of evaluating $\getbx{bx}{s}$ the first time. Then, instead of recomputing the value, the value in the table is used. The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}). Another function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).

Note as a special point that not all states are saved when evaluating a program, only when a composition is evaluated. When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, an application of $get_{mh}$, not $get$ or $get_m$, appears to indicate the saving. When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation. If there is no such key, the value $v$ will be calculated by using $get_{mh}$. The further key-value insertion is needed because the later computation could require to reevaluate $\getm{bx_1 \circ bx_2}{s}$ one more time.

>>>>

\section{Adding Memoization: minBiGUL$_m$} \label{sect:minbigulm}

When evaluating the composition of several BX programs, 
% (and the same holds for 
% full BiGUL), 
% all BX languages), 
the same $get$s are evaluated repeatedly.
% because no intermediate states are kept during the evaluation. 
This problem was illustrated in Figure~\ref{fig:eval-comp-phead}. To avoid reevaluating $get$s, and as our first approach to avoid this inefficiency, we introduce memoization in the minBiGUL interpreter.
% can save 
To keep it simple, the intermediate state of a composition is saved in a key-value table where the key is a pair of program $bx$ and source $s$, and the value is the result of evaluating $\getbx{bx}{s}$.
% the first time. 
% Then, instead of recomputing the value, the value in the table is used.
Later the value in the table is used instead of recomputing it.

The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}). The auxiliary function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).

@@

% Note as a special point 
% It should be noted 
Note in particular 
that the interpreter does not save all states when evaluating a program, only the intermediate states of a composition. When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, function $get_{mh}$ is used, not $get$ or $get_m$.
% appears to indicate the saving. 
When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation. If there is no such key, the value $v$ will be calculated by using $get_{mh}$. This key-value insertion is necessary because the interpreter may later need to reevaluate $\getm{bx_1 \circ bx_2}{s}$.
% reevaluate <--> evaluate one more time.
